---
title: "Expense Tracker for Philippines"
resource_files:
- ProjectCCA_0.0.0.9000.tar.gz
runtime: shiny
output:
  flexdashboard::flex_dashboard: null
  orientation: rows
  social: menu
  source_code: embed
---
<style>                     
.navbar {
  background-color:dark blue;
  border-color:white;
}
.navbar-brand {
color:white;
}
</style>  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(packrat)
library(PKI)
library(readxl)
library(rsconnect)
library(dplyr)
library(OutlierDetection)
library(data.table)
library(stringr)
library(shiny)
library(DT)
library(tidyverse)
library(ggplot2)
library(shinydashboard)
library(ProjectCCA)
library(gridExtra)
library(readxl)
library(dplyr)
#library(OutlierDetection)
library(data.table)
library(stringr)
library(shiny)
library(DT)
library(tidyverse)
library(ggplot2)
library(shinydashboard)
library(ProjectCCA)
#Load the data
#dt <- read.csv("C:/Users/PC-CATHERINE/Downloads/Family Income and Expenditure.csv", stringsAsFactors=FALSE)
setwd("~/Documents/Summer_Semester_2019_2020/4. Advanced Progamming in R/project/337498_425989_425148")

dt <- read.csv("Family Income and Expenditure.csv", header = TRUE) # Adinda

#dt<- filter_if(dt, is.numeric, any_vars((.) == 0))
#dt <- filter_if(dt, is.character, any_vars(!is.na(.)))
#Remove Outliers
#a<- OutlierDetection(dt[,1])
#dt$ID <- seq.int(nrow(dt))
#'%!in%' <- Negate('%in%')
#dt <- subset(dt, ID %!in% a$`Location of Outlier`)
#Change the column types
#dt <- dt %>% mutate_at(vars(matches("of|Electricity|Age|age")),as.factor)
#dt<- dt %>% mutate_if(!grepl("Expenditure|expenditure|expenses|Total|ID|Area|Value", colnames(dt)),as.factor)

#Check if the column types are correct
num_var <-colnames(select_if(dt,is.numeric))
fac_var <-colnames(select_if(dt,is.factor))



data_num <- dt %>%
  dplyr::select(num_var)

data_fac <- dt %>%
  dplyr::select(fac_var)

```


Dashboard {data-navmenu="Dynamic Charts and Table with Filters"}
===================================== 



Inputs {.sidebar}
-----------------------------------------------------------------------


```{r, echo = FALSE}

selectInput("region",
                    "Region:",
                    c(unique(as.character(dt$Region))))


selectInput("grade",
                    "Highest Grade Completed:",
                    c(unique(as.character(dt$Household.Head.Highest.Grade.Completed )))) 

selectInput("worker",
                    "Household Head Class of Worker:",
                    c(unique(as.character(dt$Household.Head.Class.of.Worker))))



```




Column 
-----------------------------------------------------------------------

### Charts

```{r, echo=FALSE}
 dt2 <- reactive({dt[dt$Region == input$region & dt$Household.Head.Highest.Grade.Completed == input$grade & dt$Household.Head.Class.of.Worker == input$worker,] %>% count(Main.Source.of.Income)})
 
 dt3 <- reactive({dt[dt$Region == input$region & dt$Household.Head.Highest.Grade.Completed == input$grade & dt$Household.Head.Class.of.Worker == input$worker,] %>% count(Household.Head.Sex)})
 
 dt4 <- reactive({dt[dt$Region == input$region & dt$Household.Head.Highest.Grade.Completed == input$grade & dt$Household.Head.Class.of.Worker == input$worker,] %>% count(Household.Head.Marital.Status)})
 
 dt5 <- reactive({dt[dt$Region == input$region & dt$Household.Head.Highest.Grade.Completed == input$grade & dt$Household.Head.Class.of.Worker == input$worker,] %>% count(Type.of.Household)})
 
 renderPlot({
     p1 <- ggplot(dt2(), aes(x= Main.Source.of.Income, y=n, fill=Main.Source.of.Income)) +geom_bar(stat="identity")+theme_minimal()
     p2 <- ggplot(dt3(), aes(x= Household.Head.Sex, y=n, fill=Household.Head.Sex)) +geom_bar(stat="identity")+theme_minimal()
     p3 <- ggplot(dt4(), aes(x= Household.Head.Marital.Status, y=n, fill=Household.Head.Marital.Status)) +geom_bar(stat="identity")+theme_minimal()
     p4 <- ggplot(dt5(), aes(x=  Type.of.Household, y=n, fill=Type.of.Household)) +geom_bar(stat="identity")+theme_minimal()
     grid.arrange(p1,p2,p3,p4, ncol=2)
   })
 
 
 
```


Row{.tabset}
-----------------------------------------------------------------------


### Summary  
```{r, echo = F} 
ui<-fluidPage(

 box(title = "Data Table", 
  status = "primary", 
  solidHeader = TRUE,
  width = 6,
  selectInput("information", label="Select Variable",
          choices = colnames(data_num), selected = "Total.Household.Income")),

box(
 title = "Data Summary", 
 status = "warning", 
 solidHeader = TRUE,
 width = 6,
 height = 142,
 verbatimTextOutput("information")))


server <-function (input,output){
 output$information <- renderPrint({
 summary(data_num[[input$information]]) 
})}

shinyApp(ui, server)
```








### Boxplot

```{r, echo=F}
#### 3.####
ui <- fluidPage(
    titlePanel("Boxplot"),
    sidebarLayout(

    # set the side panel to allow for user input
    sidebarPanel(
      selectInput(inputId="x_axis", label="Variable X", choices=colnames(data_num), selected="Total.Household.Income"),
      selectInput(inputId="y_axis", label="Variable Y", choices=colnames(data_num), selected="Agricultural.Household.indicator")
    ),
    mainPanel(plotOutput("scatterPlot"))
)
)

server <- (function(input, output) {
    output$scatterPlot <- renderPlot({
    p2 <- ggplot(data_num, aes_string(x=input$x_axis, y=input$y_axis)) + geom_boxplot()
    p2 <- p2 + xlab(input$x_axis) + ylab(input$y_axis)
    p2
  })
})


shinyApp(ui, server)
```


### Histogram

```{r, echo=F}
#### 4. ####
sliderInput("bins", "Number of bins:", min = 1, max = 50, value = 30)

renderPlot({
  x <- dt[, 1]  # Total Household income
  bins <- seq(min(x), max(x), length.out = input$bins + 1)

  # draw the histogram with the specified number of bins
  hist(x, breaks = bins, col = 'darkblue', border = 'white',
       xlab = "Household Income", main = "Histogram of Household Income",
       xlim=c(0,2000000), ylim=c(0,40000))
}, height = 500)
```


```{r, include=FALSE, echo=F}

data_2 <- dt %>% 
  count(Region) %>%
  mutate(per = round(n / sum(n) * 100)) %>%
  arrange(desc(Region))


base_plot <- data_2 %>%
  ggplot(aes("", per, fill = Region)) +
  geom_bar(stat = "identity") +
  scale_fill_discrete(name = "Region of Household") +
  xlab("") +
  ylab("Region [%]") +
  theme(plot.title = element_text(size = 30, hjust = 0.5))


min_n <- data_2[order(data_2$per),]
```

### Bar/Pie Chart

```{r, echo=F}
#### 5. ####
ui <- fluidPage(
  titlePanel("Choose a type of chart with radio buttons"),
  sidebarLayout(
  sidebarPanel(
    radioButtons("type", "Type of chart:", choices = c("Bar", "Pie")),
    textInput("title", "Tile of a chart:"),
    actionButton("accept_title", "Change title")
  ),
  mainPanel(
    plotOutput("plt")
  )
  ))

server <- function(input, output, session) {

  final_plot_wo_title <- reactive({
    plt <- base_plot
    if (input$type == "Pie") {
      plt <- plt +
        coord_polar("y") +
        geom_text(aes(x = 1, y = cumsum(per) - per / 2,
                      label = str_c(per, "%")))
    }
    plt
  })
  

  plot_title <- reactive({
    Sys.sleep(2)
    input$accept_title
    isolate(input$title)
  })
  
  
  output$plt <- renderPlot({
    final_plot_wo_title() +
      ggtitle(plot_title())
  })
}

shinyApp(ui, server)


```


### Scatter Plot

```{r, echo=F}
#### 6. ####
ui <- fluidPage(
    sidebarLayout(

    # set the side panel to allow for user input
    sidebarPanel(
      selectInput(inputId="x_axis", label="Variable X", choices=colnames(data_num), selected="Total.Household.Income"),
      selectInput(inputId="y_axis", label="Variable Y", choices=colnames(data_num), selected="Agricultural.Household.indicator")
    ),
    mainPanel(plotOutput("scatterPlot"))
)
)

server <- (function(input, output) {
    output$scatterPlot <- renderPlot({
    p1 <- ggplot(data_num, aes_string(x=input$x_axis, y=input$y_axis)) + geom_point()
    p1 <- p1 + xlab(input$x_axis) + ylab(input$y_axis)
    p1
  })
})


shinyApp(ui, server)
```

Custom Packages {data-navmenu="Custom Analysis"}
==============================================================================

Row {.tabset}
-----------------------------------------------------------------------


### AVG_Expenditure

ProjectACC packages contains 6 functions, which were written for Filipino Family Income and Expenditure dataset or any dataset having the same format. 

1.	AVG_Expenditure :  calculates the arithmetic mean of total expenditure based on a given group of region, ignoring missing values.

      AVG_Expenditure <- function(data,region){
        mydata <- data[data$Region == region,]
        AVGexpend  =  mean(mydata$Total.Food.Expenditure) + mean(mydata$Clothing..Footwear.and.Other.Wear.Expenditure)+
          mean(mydata$Imputed.House.Rental.Value)+ mean(mydata$Transportation.Expenditure) +
          mean(mydata$Education.Expenditure)+ mean(mydata$Special.Occasions.Expenditure)+
          mean(mydata$Medical.Care.Expenditure)
        return(AVGexpend)
      }

```{r, echo=FALSE, include=TRUE, echo=F}
ProjectCCA::AVG_Expenditure(dt, "CAR")
```



### Reg_Expenditure


2. Reg_Expenditure: calculates a total expenditure based on a given group of region of a given dataset

    Reg_Expenditure <- function(data, region){
      mydata <- data[data$Region == region,]
      total =  sum(mydata$Total.Food.Expenditure) + sum(mydata$Clothing..Footwear.and.Other.Wear.Expenditure)+
            sum(mydata$Imputed.House.Rental.Value) + sum(mydata$Transportation.Expenditure) +
            sum(mydata$Education.Expenditure)+ sum(mydata$Special.Occasions.Expenditure)+
            sum(mydata$Medical.Care.Expenditure)
       return(total)
    }

```{r, echo=FALSE, include=TRUE, echo=F}
ProjectCCA::Reg_Expenditure(dt, "CAR")
```




### AVG_Income


3.	AVG_Income: calculates the arithmetic income based on a given group of region, ignoring missing values.

    AVG_Income <- function(data, region){
      mydata <- data[data$Region == region,]
      avgincome <- mean(mydata$Total.Household.Income)
     return(avgincome)
    }
    
```{r, echo=FALSE, include=TRUE, echo=F}
ProjectCCA::AVG_Income(dt, "CAR")
```    



### Gender_Expenditure

4.	Gender_Expenditure: calculates a total expenditure based on a given gender

    Gender_Expenditure <- function(data,sex){
      mydata <- data[data$Household.Head.Sex == sex,]
      total =  sum(mydata$Total.Food.Expenditure) + sum(mydata$Clothing..Footwear.and.Other.Wear.Expenditure)+
        sum(mydata$Imputed.House.Rental.Value) + sum(mydata$Transportation.Expenditure) +
        sum(mydata$Education.Expenditure)+ sum(mydata$Special.Occasions.Expenditure)+
        sum(mydata$Medical.Care.Expenditure)
      return(total)
    }

```{r, echo=FALSE, include=TRUE, echo=F}
renderPlot( ProjectCCA::ExpenditurePlot(dt, "Household.Head.Sex" ), width = 1900, height = 700)
```



### ExpenditurePlot


5.	ExpenditurePlot :  returns a plot of how expenditure depends on a given variable of a given data. If the variables is factor, it returns a boxplot, it the variables is numeric, it returns scatter plot.

    ExpenditurePlot <- function(data, var){
      varx <- data[, var]
      
      Expenditure <- data$Total.Food.Expenditure + data$Clothing..Footwear.and.Other.Wear.Expenditure +
        data$Imputed.House.Rental.Value + data$Transportation.Expenditure +
        data$Education.Expenditure + data$Special.Occasions.Expenditure +
        data$Medical.Care.Expenditure
      
      if(is.factor(varx)) {
        boxplot(Expenditure ~ varx,
                data = data,
                xlab = var,
                ylab = "Total Expenditure")
      } else if ( is.numeric(varx)){
        plot(x = varx, y = Expenditure,
             xlab = var,
             ylab = "Total Expenditure")
      } else {
        print("No plot")
      }
    }

```{r, echo=FALSE, include=TRUE, echo=F}
renderPlot( ProjectCCA::ExpenditurePlot(dt, "Household.Head.Sex" ), width = 2000, height = 700)
```


### my_reg


6.	my_reg: returns a regression plot and p-values for a given formula based on a given data

    my_reg <- function(formula, data) {
      fit = lm(formula, data)
      par(mfrow=c(1,1)) 
      plot(fit)
      print(summary(fit))
      invisible(fit)
    }

```{r, echo=FALSE, include=TRUE, echo=F}
ProjectCCA::my_reg(Housing.and.water.Expenditure ~ Number.of.Refrigerator.Freezer + Number.of.Airconditioner + Number.of.Landline.wireless.telephones +
                     Number.of.Personal.Computer + Number.of.Television + Number.of.Washing.Machine + House.Floor.Area, dt )

```



Table {data-navmenu="Dynamic Charts and Table with Filters"}
===================================== 


Row {.tabset}
-----------------------------------------------------------------------

### Table with specific Filters


```{r, echo = FALSE}
#### 8. ####
ui <- fluidPage(
  titlePanel("Basic DataTable"),

  # Create a new Row in the UI for selectInputs
 
   fluidRow(
    column(4,
        selectInput("region",
                    "Region:",
                    c("All",
                      unique(as.character(dt$Region))))
    ),
    column(4,
        selectInput("grade",
                    "Highest Grade Completed:",
                    c("All",
                      unique(as.character(dt$Household.Head.Highest.Grade.Completed ))))
    ),
    column(4,
        selectInput("worker",
                    "Household Head Class of Worker:",
                    c("All",
                      unique(as.character(dt$Household.Head.Class.of.Worker))))
    )
  ),
  # Create a new row for the table.
  DT::dataTableOutput("table")
)


server <- function(input, output) {

  # Filter data based on selections
  output$table <- DT::renderDataTable(DT::datatable({
    data <- dt
    if (input$region != "All") {
      data <- data[data$Region == input$region,]
    }
    if (input$grade != "All") {
      data <- data[data$Household.Head.Highest.Grade.Completed == input$grade,]
    }
    if (input$worker != "All") {
      data <- data[data$Household.Head.Class.of.Worker == input$worker,]
    }
    data
  }),class = "display nowrap compact",
    filter = "top")

}

shinyApp(ui, server)
```

### Table with All Filters

```{r, echo=F}
#### 1. Showing data ####
ui <- fluidPage(fluidRow
                (column(12, div(dataTableOutput("dataTable")))))


server <- function(input, output, session) {
  output$dataTable <- renderDT(
    dt, 
    class = "display nowrap compact",
    filter = "top" 
  )

}

shinyApp(ui, server)
```
